---
title: "priceDiscoveRy: Package Overview and Methodology"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
author: "Hui Sun"
vignette: >
  %\VignetteIndexEntry{priceDiscoveRy: Package Overview and Methodology}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette first introduces the scope and interface of `priceDiscoveRy` and
then presents the economic background, assumptions, and methodology behind the
package. The methodological discussion follows the source-file order used by
the package so that each theoretical object is linked to the function that
implements it. To avoid repetition, each estimator is defined only once; later
sections refer back to that definition.

# The `priceDiscoveRy` package

## Purpose and scope

`priceDiscoveRy` estimates contribution-weighted information shares (CWIS) for
two related markets with partially overlapping trading hours. The current
implementation is written for a futures market and a spot market that represent
the same asset, or economically equivalent assets linked by a stable long-run
equilibrium relation.

Conventional Hasbrouck information shares (HIS) describe price discovery while
both markets trade. They do not measure the information incorporated while only
one market remains open. CWIS extends the analysis to the complete trading day
by combining:

1. the HIS midpoint estimated during overlapping trading;
2. the variation in the common efficient price generated in overlapping
   subperiods; and
3. the variation generated in futures-only and spot-only subperiods.

The periods are weighted by their estimated contributions to daily
efficient-price variation, not by their clock-time duration. Consequently, a
short but informative session can receive more weight than a long, quiet
session.

## User interface and result objects

The main user-level function is `cwis()`. It validates the data, splits the
sample by day, performs the complete calculation for every usable day, and
aggregates the results. A typical call has the following form:

```{r main-interface}
fit <- cwis(
  data = market_data,
  datetime_col = "datetime",
  futures_col = "futures_price",
  spot_col = "spot_price",
  spot_multiplier = 10,
  tz = "America/New_York",
  K = 10L,
  kernel_type = "ModifiedTukeyHanning",
  align_by = "seconds",
  align_period = 1L,
  continue_on_error = TRUE
)
```

The returned `cwis_result` contains daily estimates, cross-day summaries,
diagnostics, average variance weights, detailed daily objects, and the settings
used in the calculation. Successful daily objects retain the lower and upper
HIS bounds, their midpoint, session variance estimates and weights, CWIS, VECM
residual correlation, zero-return fractions, and observation counts. Failed
days remain visible when `continue_on_error = TRUE`; they should be investigated
rather than silently discarded.

## Function map in source order

The source files form one computational pipeline. The order below matches the
numbered files in the package.

| Order | Source file | Main function | Role |
|:--:|:--|:--|:--|
| 1 | `1_validate_and_standardize_input.R` | `validate_and_standardize_input()` | Parse, scale, validate, sort, and standardize the two price series. |
| 2 | `2_split_trading_days.R` | `split_trading_days()` | Split standardized observations into local calendar days. |
| 3 | `3_split_trading_sessions.R` | `split_trading_sessions()` | Allocate a day to overlapping and single-market sessions. |
| 4 | `4_vecm.R` | `vecm()` | Estimate the overlapping-period Johansen VECM. |
| 5 | `5_common_price.R` | `common_price()` | Recover the permanent common-price component. |
| 6 | `6_his.R` | `his()` | Compute ordering-specific Hasbrouck information shares. |
| 7 | `7_realized_variance.R` | `realized_variance()` | Estimate variation in common-price returns during overlap. |
| 8 | `8_realized_kernel.R` | `realized_kernel()` | Estimate noise-robust variation in single-market sessions. |
| 9 | `9_combine_weights.R` | `combine_weights()` | Normalize session variances into contribution weights. |
| 10 | `10_daily_cwis.R` | `daily_cwis()` | Combine the components into one daily CWIS result. |
| 11 | `11_aggregate_cwis.R` | `aggregate_cwis()` | Summarize successful and failed days. |
| 12 | `12_cwis.R` | `cwis()` | Provide the complete multi-day interface. |
| 13 | `13_plot_cwis.R` | `plot.cwis_result()` | Compare daily HIS and CWIS graphically. |

# Background

## Price discovery in linked markets

The same or an economically equivalent asset may trade in more than one venue.
Observed prices can differ temporarily because of bid-ask bounce, discreteness,
latency, liquidity, and other market-microstructure effects. An arbitrage or
cost-of-carry relationship nevertheless prevents the two prices from drifting
apart indefinitely. Their non-stationary components therefore reflect a common
efficient price.

Hasbrouck's (1995) information share attributes innovations in that common
efficient price to the observed markets. It captures both how quickly a market
incorporates new information and how well its price avoids transitory noise.
The standard measure requires simultaneous observations, so it describes
leadership conditional on both markets being open.

In the latent-price model used by Dimpfl and Schweikert (2023), the unobserved
efficient price follows a random walk,

\[
m_{d,t}=m_{d,t-1}+e_{d,t},
\]

and an observed price in open market (j) satisfies

\[
p_{j,t}=m_{d,t}+\varepsilon_{j,t}.
\]

The efficient-price innovation (e_{d,t}) represents permanent information,
whereas (\varepsilon_{j,t}) represents transitory market-microstructure
frictions. This distinction motivates using the VECM common trend to measure
overlap-period variation and a noise-robust realized kernel when only one venue
is observed.

Partially overlapping hours add a second dimension. Information may arrive
while only one market trades. Restricting the sample to the overlap omits this
part of the daily discovery process, whereas weighting periods by duration
would treat quiet and informative periods alike. Dimpfl and Schweikert (2023)
resolve the problem by weighting each session by its contribution to daily
efficient-price variation. During a single-market session, the open market is
assigned the information generated in that session; during overlap, the
allocation is determined by HIS.

## Partitioning the trading day

Let day (d) contain (T) ordered observation intervals and let
(S_1,\ldots,S_H) be disjoint subperiods whose union is the complete day. A
subperiod may contain simultaneous trading or trading in only one market. The
partition is economic, not merely chronological: it changes whenever the set
of open markets changes and may also separate economically distinct parts of
the overlap, such as pre-core, core, and post-core trading.

With the package defaults, the day is divided in New York local time as follows:

| Component session | Interval | Open market(s) |
|:--|:--|:--|
| `futures_pre` | before 04:00:00 | futures only |
| `overlap_pre` | 04:00:00 to 09:30:00 | futures and spot |
| `overlap_core` | 09:30:00 to 16:00:01 | futures and spot |
| `overlap_post_1` | 16:00:01 to 17:00:00 | futures and spot |
| `spot_maintenance` | 17:00:00 to 18:00:00 | spot only |
| `overlap_post_2` | 18:00:00 to 20:00:00 | futures and spot |
| `futures_post` | from 20:00:00 | futures only |

Except for the first and last components, the intervals include the lower and
exclude the upper boundary. The default `core_close_exclusive` value of
16:00:01 was designed for one-second data so that 16:00:00 remains in the core
session. Sub-second applications should replace this boundary with a value
appropriate to their sampling grid.

# Assumptions and empirical scope

The assumptions below collect the conditions used by the theoretical framework
and the additional requirements imposed by the implementation.

## Economic and statistical assumptions

1. **Common fundamental value.** The two prices refer to the same asset or to
   assets connected by a stable arbitrage relation.
2. **One permanent trend.** During overlapping trading, the two log prices are
   individually non-stationary, cointegrated with rank one, and driven by one
   common stochastic trend.
3. **Efficient-price evolution.** The latent efficient price continues to
   evolve whether one, both, or neither observed venue is open. Its increments
   are independent and mean zero, but need not be identically distributed;
   their variance may be heteroskedastic or exhibit deterministic intraday
   seasonality.
4. **Transitory observation error.** An observed open-market price equals the
   efficient price plus a mean-zero, weakly dependent, covariance-stationary
   microstructure component. Its variance may change over time and differ
   between markets.
5. **Law of one price during overlap.** Short-run deviations are allowed, but
   the two simultaneously observed prices share the same long-run value.
6. **Stable overlapping-period dynamics.** The VECM parameters are treated as
   stable across the overlapping observations pooled within a trading day.
7. **One-market attribution.** During a single-market session, the open market
   receives a conditional information share of one and the closed market a
   share of zero.
8. **Comparable variance concept.** All session estimators measure the same
   efficient-price variation in squared log-return units, so they can be added
   and normalized.
9. **Cholesky uncertainty.** When contemporaneous VECM innovations are
   correlated, their covariance cannot be uniquely assigned to markets. Both
   orderings are therefore retained as bounds and summarized by their midpoint.

## Data and implementation requirements

- Timestamps must identify unique observation instants and be interpreted in
  the correct IANA time zone.
- Prices must be numeric, strictly positive when observed, and placed on
  economically comparable scales before taking logarithms.
- Missing values may identify a closed market, but the two prices must both be
  observed in overlapping sessions and the open-market price must be observed
  in each single-market session.
- Each of the seven required component sessions must contain at least two rows;
  the pooled overlap must contain more than `K + 2` complete observations and
  both price series must vary.
- The sampling frequency, session boundaries, VECM lag order, kernel alignment,
  and bandwidth must be chosen for the market and data source rather than
  accepted mechanically.
- Local-calendar-day splitting is suitable only when the economic trading day
  agrees with the chosen local date. Markets whose trading day begins on the
  preceding evening require a custom day convention upstream.

## Relationship to the reference methodology

The package implements the CWIS framework of Dimpfl and Schweikert (2023), but
it is not an exact replication of their high-resolution empirical estimator.
The paper uses a HAR-structured VECM for 10-millisecond observations, with lag
blocks extending to ten seconds. `priceDiscoveRy` instead calls
`urca::ca.jo()` to fit a standard finite-lag Johansen VECM. The variance
weighting, HIS bounds, common-trend extraction, realized-variance calculation,
realized-kernel calculation, and final CWIS combination follow the same
conceptual framework, while the short-run VECM parameterization differs.

# Methodology in package order

## 1. Validate and standardize the input

`validate_and_standardize_input()` maps the user-selected columns to the
internal schema `datetime`, `trading_day`, `futures`, and `spot`. Character
timestamps are parsed in the supplied time zone, observations are sorted, and
duplicate timestamps are rejected. The spot series is multiplied by
`spot_multiplier` (default 10) before validation so that the two prices can be
placed on comparable scales.

This step does not impute prices. It allows `NA` values that represent market
closures, but every observed value must be finite and strictly positive. The
local date derived from the standardized timestamp becomes `trading_day`.

## 2. Split observations into trading days

`split_trading_days()` separates the standardized data by `trading_day` and
preserves the time-zone attribute. Every later model is therefore estimated
independently for one local calendar day. This design permits day-specific
information shares and variance weights rather than imposing a single estimate
on the full sample.

## 3. Split each day into trading sessions

`split_trading_sessions()` assigns each observation to the seven component
sessions defined above and joins the four simultaneous-trading components into
`overlap`. The function uses local clock time rather than row positions. Custom
boundaries must retain the six required names and be strictly increasing within
the 24-hour day.

The pooled `overlap` sample supplies the VECM, while its component-session row
counts later allocate common-price returns to economically meaningful
subperiods.

## 4. Estimate the overlapping-period VECM

Let (p_t=(p_{F,t},p_{S,t})') contain the futures and spot log prices observed
during overlap. The standard VECM used by `vecm()` is

\[
\Delta p_t
= \Pi p_{t-1}
+ \sum_{i=1}^{K-1}\Gamma_i\Delta p_{t-i}
+ d_t + u_t,
\qquad
\operatorname{Var}(u_t)=\Omega,
\tag{1}
\]

where (\Pi=\alpha\beta') contains the long-run relation and adjustment
coefficients, (\Gamma_i) describes short-run dynamics, (d_t) contains any
deterministic terms, and (u_t) is the reduced-form innovation vector. The
package default is a Johansen trace specification with `ecdet = "none"`,
`spec = "transitory"`, and `K = 10`. In `urca::ca.jo()`, `K` is the lag order
in levels, so the VECM contains (K-1) lags of first differences.

The daily estimator assumes rank one downstream. The Johansen object is
retained, but `daily_cwis()` does not perform a separate data-driven rank choice
for every day; users should verify cointegration and the economic restriction
before interpreting the results.

## 5. Extract the common efficient price

For rank one, let (\alpha_\perp) and (\beta_\perp) denote orthogonal
complements and define

\[
\Upsilon = I_2-\sum_{i=1}^{K-1}\Gamma_i,
\qquad
C = \beta_\perp
\left(\alpha_\perp'\Upsilon\beta_\perp\right)^{-1}
\alpha_\perp'.
\tag{2}
\]

The common-trend representation separates the permanent response
(C\sum_{s=1}^{t}u_s) from stationary transitory dynamics. Because (C) has
rank one, its rows describe the same permanent trend up to normalization.
`common_price()` returns the first component of this cumulative permanent
part, the long-run impact matrix (C), and the contemporaneous correlation of
the two VECM residual series.

The residual correlation is diagnostically important: stronger correlation
usually makes the ordering-specific HIS estimates farther apart.

## 6. Calculate Hasbrouck information shares

Let (\xi) be a row of (C) and write the innovation covariance matrix as
(\Omega=FF'), where (F) is a lower-triangular Cholesky factor. For the
supplied market ordering, `his()` calculates

\[
S_j = \frac{[\xi F]_j^2}{\xi\Omega\xi'},
\qquad j\in\{F,S\}.
\tag{3}
\]

When the innovations are correlated, the Cholesky allocation depends on which
market is placed first. `daily_cwis()` therefore fits the VECM twice: once in
futures-spot order and once in spot-futures order. After restoring common market
labels, the two estimates define each market's lower and upper bound. The
overlapping-period point estimate is their midpoint,

\[
S_j^* = \frac{S_j^{\mathrm{lower}}+S_j^{\mathrm{upper}}}{2}.
\tag{4}
\]

The midpoint is a summary, not a resolution of the identification problem. The
bounds and residual correlation should remain part of the reported evidence.

## 7. Estimate overlapping-period realized variance

Let (\Delta \widehat m_i) denote returns of the estimated common price. For an
overlapping subperiod (S_h), `realized_variance()` uses

\[
\widehat{IV}_{h}^{\mathrm{RV}}
= \sum_{i\in S_h}(\Delta\widehat m_i)^2.
\tag{5}
\]

The common-price innovations are serially uncorrelated by construction under
the maintained VECM representation, which motivates ordinary realized variance
rather than a univariate venue-price estimator during fragmented trading.

The implementation reports three overlap variance components:
`overlap_pre`, `overlap_core`, and `overlap_post`. The last component combines
the row counts of `overlap_post_1` and `overlap_post_2`. The complete pooled
overlap is used to estimate one VECM before its efficient returns are allocated
to these components.

## 8. Estimate single-market realized kernels

Observed returns in a single open market may be autocorrelated by
microstructure noise. For each futures-only or spot-only session,
`realized_kernel()` estimates

\[
\widehat{IV}_{h}^{K}
= \sum_{i=1}^{n}r_{i,n}^{2}
+ \sum_{b=1}^{B}
k\!\left(\frac{b-1}{B}\right)
(\eta_b+\eta_{-b}),
\qquad
\eta_b=\sum_{i=1}^{n-b}r_{i,n}r_{i+b,n}.
\tag{6}
\]

The weighted autocovariances correct for serial dependence at high frequency.
The default is the modified Tukey-Hanning kernel used in the reference paper.
The wrapper sets

\[
B=\left\lfloor c\sqrt{N}\right\rfloor,
\tag{7}
\]

where (N) is the number of session rows and (c=2.3970) for the modified
Tukey-Hanning kernel or (c=3.5134) for the Parzen kernel. Other kernels require
an explicit positive `bandwidth_constant`. `align_by` and `align_period`
control sampling before returns are constructed; `align_by = NULL` retains an
already prepared observation grid.

The three single-market estimates are `futures_pre`, `spot_maintenance`, and
`futures_post`. They are expressed in squared-return units and are neither
annualized nor converted to volatility.

## 9. Convert session variances into weights

Let (\widehat\sigma_{m,h}^{2}) be the estimated efficient-price variance for
period (S_h), obtained from Equation (5) or (6) as appropriate.
`combine_weights()` calculates

\[
w_h =
\frac{\widehat\sigma_{m,h}^{2}}
{\sum_{q=1}^{H}\widehat\sigma_{m,q}^{2}},
\qquad
\sum_{h=1}^{H}w_h=1.
\tag{8}
\]

The total overlap weight is the sum of `overlap_pre`, `overlap_core`, and
`overlap_post`. The futures-only weight is the sum of `futures_pre` and
`futures_post`; the spot-only weight is `spot_maintenance`.

These are contribution weights, not precision weights. They deliberately have
a direct variance-share interpretation and do not adjust for possible
differences in sampling error or estimator accuracy across sessions.

## 10. Construct daily CWIS

`daily_cwis()` orchestrates Steps 3-9. Write (w_O) for the total overlap
weight, (w_F) for the futures-only weight, and (w_S) for the spot-only
weight. With (S_F^*) and (S_S^*) denoting the overlapping-period HIS
midpoints, the daily measure is

\[
\begin{pmatrix}
CWIS_F\\
CWIS_S
\end{pmatrix}
=
w_O
\begin{pmatrix}
S_F^*\\
S_S^*
\end{pmatrix}
+
w_F
\begin{pmatrix}
1\\
0
\end{pmatrix}
+
w_S
\begin{pmatrix}
0\\
1
\end{pmatrix}.
\tag{9}
\]

This is the only stage at which the conditional information shares and the
across-period variance weights are combined. The identities
(w_O+w_F+w_S=1), (S_F^*+S_S^*=1), and
(CWIS_F+CWIS_S=1) should hold up to numerical precision.

Two limiting cases clarify the interpretation. If (w_O=1), daily CWIS equals
the conventional HIS midpoint. If there is no overlap, or an overlapping
market has a conditional share of one, the expression reduces to the
sequential-market allocation when the same variance estimator is used.

## 11. Aggregate across days

`aggregate_cwis()` converts daily objects and failure records into a single
daily table. Across successful days it reports the arithmetic mean and sample
standard deviation of the HIS midpoint and CWIS, along with mean and standard
deviation of each period weight. It also reports counts of supplied,
successful, and failed days, the mean and standard deviation of the overlap
weight, and mean residual correlation.

The aggregation is descriptive and gives every successful day equal weight. It
does not re-estimate a pooled VECM, pool intraday variances across days, or
weight daily CWIS by the day's total variance.

## 12. Run the complete multi-day workflow

`cwis()` connects the preceding stages in the same order as the source files:

\[
\text{raw data}
\rightarrow \text{standardized days}
\rightarrow \text{sessions}
\rightarrow \{\text{VECM, HIS, common price, RV, RK}\}
\rightarrow \text{weights}
\rightarrow \text{daily CWIS}
\rightarrow \text{cross-day summary}.
\]

When `continue_on_error = TRUE`, a failed day contributes an error record but
does not stop later dates from being processed. This behavior supports long
samples, but the reason and frequency of failure remain part of the empirical
result.

## 13. Plot HIS and CWIS

`plot.cwis_result()` is an interpretive endpoint rather than an additional
estimator. It sorts successful days, plots the selected market's HIS midpoint
and CWIS, and adds a 0.5 benchmark by default. Failed days are omitted from the
plot but remain in `fit$daily`; therefore the plot should always be read with
the failure diagnostics.

# Interpretation and diagnostics

HIS and CWIS answer related but different questions:

- HIS asks which market leads price discovery while both markets trade.
- CWIS asks how much each market contributes over the complete trading day.

A difference between the two does not mean that the overlapping-period VECM
changed. It means that economically relevant variation occurred when only one
market was open. The session weights identify where that difference originated.

For every application, interpretation should proceed in the following order:

1. verify the time zone, trading-day convention, market-open indicators, price
   scaling, and session boundaries;
2. examine failed days and confirm that exclusions are not systematic;
3. verify cointegration rank one and assess the chosen VECM lag order;
4. inspect HIS lower and upper bounds together with residual correlation;
5. inspect zero-return fractions for stale prices or an excessively fine grid;
6. inspect all session variance weights and confirm that they sum to one;
7. compare HIS with CWIS and attribute the difference to specific
   single-market weights; and
8. check robustness to plausible sampling frequencies, kernel alignment,
   bandwidths, and session definitions.

CWIS is a variance-based measure of permanent information incorporation. It is
not, by itself, a measure of trading volume, liquidity, profitability, causal
influence, or overall market quality. The package reports daily dispersion but
does not currently provide analytical standard errors or confidence intervals
for HIS, session weights, or CWIS.

# References

Barndorff-Nielsen, O. E., Hansen, P. R., Lunde, A., and Shephard, N. (2008).
Designing realized kernels to measure the ex post variation of equity prices in
the presence of noise. *Econometrica*, 76(6), 1481-1536.
<https://doi.org/10.3982/ECTA6495>.

Barndorff-Nielsen, O. E., Hansen, P. R., Lunde, A., and Shephard, N. (2009).
Realized kernels in practice: Trades and quotes. *The Econometrics Journal*,
12(3), C1-C32. <https://doi.org/10.1111/j.1368-423X.2008.00275.x>.

Dias, G. F., and Schweikert, K. (2022). Integrated variance estimation for
assets traded in multiple venues. *SSRN Working Paper*.
<https://ssrn.com/abstract=4253762>.

Dimpfl, T., and Schweikert, K. (2023). Information shares for markets with
partially overlapping trading hours. *Journal of Banking & Finance*, 154,
106970. <https://doi.org/10.1016/j.jbankfin.2023.106970>.

Hasbrouck, J. (1995). One security, many markets: Determining the contributions
to price discovery. *The Journal of Finance*, 50(4), 1175-1199.
<https://doi.org/10.2307/2329348>.

Hasbrouck, J. (2021). Price discovery in high resolution. *Journal of Financial
Econometrics*, 19(3), 395-430. <https://doi.org/10.1093/jjfinec/nbz027>.

Stock, J. H., and Watson, M. W. (1988). Testing for common trends. *Journal of
the American Statistical Association*, 83(404), 1097-1107.
<https://doi.org/10.2307/2290142>.

Wang, J., and Yang, M. (2011). Housewives of Tokyo versus the gnomes of Zurich:
Measuring price discovery in sequential markets. *Journal of Financial
Markets*, 14(1), 82-108. <https://doi.org/10.1016/j.finmar.2010.08.002>.

